File-based localization of numerical perturbations in data analysis pipelines

### 2020-08-21 paper.tex
Update revisions

\usepackage{subcaption}
\usepackage{xspace}
\usepackage{stmaryrd} % for llbracket and rrbracket
\usepackage{amsmath}
\usepackage{ulem}
\usepackage{pifont}

@@

\revised{
The labeled graph can differ depending on the order of executions in which condition we capture transient files 
or execute the pipeline to pinpoint the propagation of differences. Therefore, we label processes in two orders 
of execution, and then aggregate findings so that label processes as non-reproducible (red) if it creates 
different results in either order of executions.}

\subsection{Implementation}
\revised{
\toolname is implemented in Python (>=3.5). In this work we used \toolname version 0.2 and the following version of
the Python package dependencies: NumPy v1.19.0~\cite{oliphant2006guide} and Pandas v1.0.5~\cite{mckinney2011pandas},
for data manipulations, SciPy v1.5.1~\cite{2020SciPy-NMeth} and Scikit-learn v.0.23.1~\cite{pedregosa2011scikit}
for the clustering of provenance graphs, Zss v1.2.0~\cite{zhang1989simple} for tree distances,
\reprozip v1.0.11 for the capture of provenance traces, Docker v17.05~\cite{merkel2014docker} for
the edition of container images, and Boutiques v0.5.25~\cite{glatard2018boutiques} for uniform pipeline executions.

@@

% \revision{Using the \reprozip trace obtained from CentOS7,
% we captured transient files in CentOS6 and labeled pipeline processes in CentOS7.}
We processed the 20 subjects with PreFreeSurfer and FreeSurfer, using the 2 CentOS versions.
\revised{First, PreFreeSurfer was utilized to analyze the unprocessed data, and then its results 
from CentOS6 were used as the input of FreeSurfer in both conditions. 
We also used the ReproZip trace file captured in CentOS6 for labeling the processes in both pipelines.}
Each subject was processed twice on the same operating
system to detect within-OS variability coming from pseudo-random
operations. We compared pipeline results using  FreeSurfer tools \texttt{mri\_diff},
\texttt{mris\_diff}, and \texttt{lta\_diff}, to ignore execution-specific information such as file path or
timestamps. To compare segmentations $X$ and $Y$, we used the Dice coefficient defined as follows:
\[DICE=\frac{2|X \cap Y|}{|X| + |Y|}\]

\revised{
The Dice coefficient~\cite{dice1945measures} is a commonly used metric to validate medical image segmentation.
Dice values range from 0 to 1, with 1 indicating a perfect overlap between two segmentation results and 0 indicates no overlap.
Alternatively, the Jaccard coefficient~\cite{jaccard1912distribution} could be used;
there is a direct correspondence between both metrics.}

@@

\section{Results}
\revised{
All experiments were run on a machine with a 3.4GHz Intel Core i7 processor using 8 cores and 32GB of RAM running
the CentOS7 operating system and Linux kernel version 3.10.
The processing time, output file size, number of file accesses and number of processes observed in PreFreeSurfer
and FreeSurfer are shown in~Table~\ref{table:pipeline-stats}.
The scripts and analyses used to create the figures in this section are available at
\url{https://github.com/big-data-lab-team/HCP-reproducibility-paper}.

@@

   &        \makecell{Mean}   &  \makecell{Standard error}  & \makecell{Mean}   &  \makecell{Standard error}   \\ \midrule
  Processing time (mins)      &     106.67     &     2.68     &     nd     &     nd     \\
  Output file size (GB)       &     2.8        &     0.10     &     nd     &     nd     \\
  Number of file accesses     &     94,089      &     2,645  &     nd     &     nd     \\
  Number of processes         &     8731       &     198   &     nd     &     nd     \\
  \bottomrule
  \end{tabular}
  \end{threeparttable}
  \end{table*}
}
% All experiments were run on a machine with a 3.4GHz Intel Core i7 processor
% and 32GB memory under the CentOS7 operating system.
% The average processing time per subject was approximately 2~hours (with standard error X)
% for PreFreeSurfer and 8~hours (with standard error X) for FreeSurfer.
% The average output file size was 2.7~GB (with standard error Y) for PreFreeSurfer
% and 4.1~GB (with standard error Y) for FreeSurfer.
%\tristan{Ali, are these numbers per subject or for 20 subjects? it's an average of 20 subjects}
% For each subject, PreFreeSurfer accessed 83K files and created 7.7K processes, and FreeSurfer accessed 62K files and
% created 4K processes.

@@

\revised{
Figure~\ref{fig:fnirt_result} compares \fnirt results in Brain Extraction
for a particular subject using the checkerboard pattern,
a common method to illustrate the magnitude of the differences in registration results.
Differences appear to be visually important, in particular in the areas framed in red.}

@@

However, FreeSurfer results still differ between conditions, due to the
propagation of differences created in PreFreeSurfer. We
observed the effect of this propagation in FreeSurfer results, as shown in
Figure~\ref{fig:tissue_class} for whole-brain segmentations. The
Dice coefficients associated with the 44 regions segmented by
FreeSurfer are shown in Figure~\ref{fig:scatter_plot}, showing that Dice coefficents below 0.9 are
observed in most regions, and particularly in the smallest ones.
\revised{However, no significant correlation between the Dice values and the region sizes was found
(Pearson's coefficient = 0.12, p-value = 0.43).}

@@

\revised{
% Furthermore, we can use methods to ensure that enough tests have been performed for labeling a tool as reproducible.
% We approved the irreproducibility of the tools, but we cannot say a tool is reproducible since
\toolname labels pipeline processes based on the tested dataset, 
while the reproducibility of the tools can be changed over the new datasets. 
In this regard, we need to confirm the reproducibility of the tools over all the possible 
subject diversity and processing parameters. 
We can rely on the code coverage to measure the proportion of the code covered by the tests. 
Ideally, the code coverage as close to 100\% as possible ensures that most of the code is fully tested. 
This is a key principle of reproducibility to guarantee the appropriate testing of the tools. 
Having sufficient subject tests so that covers most of the code could increase confidence to confirm 
the reproducibility of the tools.
As another solution, we can estimate the probability of the tool reproducibility based on 
the number of subjects that the tool was tested as reproducible. To this order, 
we refer to the probability measurement used in~\cite{glatard2015classifications}, which can state that the probability 
to observe all reproducible results among $N’$ future tests given that $n$ non-reproducible results happened among $N$ 
previous tests is:

\[p(N',n,N)=\frac{N+1}{N+N'+1}\frac{\binom Nn}{\binom {N+N'}n}\]

This formula gives a high probability to see reproducible results from the tools have been tested on a high series of 
subjects as reproducible.
}

@@

\revised{
\section{Availability of Source Code and Requirements}
\begin{itemize}
  \item Project name: \toolname
  \item Project home page:\url{https://github.com/big-data-lab-team/spot}
  \item Operating system: Linux
  \item Programming language: Python
  \item Other requirements: Python 3.6 or higher, \reprozip, Docker, and Boutiques
  \item License: MIT License
  \item Biotools identifier: \href{https://bio.tools/spottool}{spottool}
  \item SciCrunch ID: RRID:SCR\_018915

@@

>>>>

\usepackage{subcaption}
\usepackage{xspace}
\usepackage{stmaryrd} % for llbracket and rrbracket
\usepackage{amsmath}
\usepackage{ulem}
\usepackage{pifont}
\usepackage{pdfpages}

@@

\revised{
The labeled graph can differ depending on the order of executions in which condition we capture transient files 
or execute the pipeline to pinpoint the propagation of differences. Therefore, we run the comparison in both condition orders, 
and we label a process as non-reproducible (red) if it creates different results in at least one condition order.}

\subsection{Implementation}
\revised{
\toolname is implemented in Python (>=3.6). In this work we used \toolname version 0.2 and the following version of
the Python package dependencies: NumPy v1.19.0~\cite{oliphant2006guide} and Pandas v1.0.5~\cite{mckinney2011pandas},
for data manipulations, SciPy v1.5.1~\cite{2020SciPy-NMeth} and Scikit-learn v.0.23.1~\cite{pedregosa2011scikit}
for the clustering of provenance graphs, Zss v1.2.0~\cite{zhang1989simple} for tree distances,
\reprozip v1.0.11 for the capture of provenance traces, Docker v17.05~\cite{merkel2014docker} for
the edition of container images, and Boutiques v0.5.25~\cite{glatard2018boutiques} for uniform pipeline executions.

@@

% \revision{Using the \reprozip trace obtained from CentOS7,
% we captured transient files in CentOS6 and labeled pipeline processes in CentOS7.}
We processed the 20 subjects with PreFreeSurfer and FreeSurfer, using the 2 CentOS versions.
\revised{The PreFreesurfer results obtained in CentOS6 were used as the input of FreeSurfer in both conditions. 
We also used the \reprozip trace file captured in CentOS6 for labeling the processes in both pipelines.}
Each subject was processed twice on the same operating
system to detect within-OS variability coming from pseudo-random
operations. We compared pipeline results using  FreeSurfer tools \texttt{mri\_diff},
\texttt{mris\_diff}, and \texttt{lta\_diff}, to ignore execution-specific information such as file path or
timestamps. To compare segmentations $X$ and $Y$, we used the Dice coefficient defined as follows:
\[DICE=\frac{2|X \cap Y|}{|X| + |Y|}\]

\revised{
The Dice coefficient~\cite{dice1945measures} is a commonly used metric to validate medical image segmentation.
Dice values range from 0 to 1, with 1 indicating a perfect overlap between two segmentation results and 0 indicating no overlap.
Alternatively, the Jaccard coefficient~\cite{jaccard1912distribution} could be used;
there is a direct correspondence between both metrics.}

@@

\section{Results}
\revised{
All experiments were run on a machine with a 3.4GHz, 8-core  Intel Core i7 processor, 32GB of RAM, 
CentOS 7.3.1611, and Linux kernel version 3.10.
The processing time, output file size, number of file accesses and number of processes observed in PreFreeSurfer
and FreeSurfer are shown in~Table~\ref{table:pipeline-stats}.
The scripts and analyses used to create the figures in this section are available at
\url{https://github.com/big-data-lab-team/HCP-reproducibility-paper}.

@@

                          & \makecell{Mean}  &  \makecell{Standard error}  & \makecell{Mean}   &  \makecell{Standard error}   \\ \midrule
  Processing time (mins)  &     106.67       &     2.68                    &     650.25        &     8.88     \\
  Output file size (GB)   &     2.8          &     0.10                    &     4.15          &     0.15     \\
  Number of file accesses &     94,089       &     2,645                   &     62,729        &     984     \\
  Number of processes     &     8,731         &     198                     &     4,031          &     47     \\
  \bottomrule
  \end{tabular}
  \end{threeparttable}
  \end{table*}
}
% All experiments were run on a machine with a 3.4GHz Intel Core i7 processor
% and 32GB memory under the CentOS7 operating system.
% The average processing time per subject was approximately 2~hours (with standard error X)
% for PreFreeSurfer and 8~hours (with standard error X) for FreeSurfer.
% The average output file size was 2.7~GB (with standard error Y) for PreFreeSurfer
% and 4.1~GB (with standard error Y) for FreeSurfer.
%\tristan{Ali, are these numbers per subject or for 20 subjects? it's an average of 20 subjects}
% For each subject, PreFreeSurfer accessed 83K files and created 7.7K processes, and FreeSurfer accessed 62K files and
% created 4K processes.

@@

\revised{
Figure~\ref{fig:fnirt_result} compares \fnirt results in Brain Extraction
for a particular subject using the checkerboard pattern,
a common method to illustrate the magnitude of the differences in registration results.
Differences appear to be visually important, in particular in the areas framed in red, 
to the point that most experimenters would likely reject such a registration following visual quality control.}

@@

However, FreeSurfer results still differ between conditions, due to the
propagation of differences created in PreFreeSurfer. We
observed the effect of this propagation in FreeSurfer results, as shown in
Figure~\ref{fig:tissue_class} for whole-brain segmentations. The
Dice coefficients associated with the 44 regions segmented by
FreeSurfer are shown in Figure~\ref{fig:scatter_plot}, showing that Dice coefficients below 0.9 are
observed in most regions, and particularly in the smallest ones.
\revised{However, no significant correlation between the Dice values and the region sizes was found
(Pearson's coefficient = 0.12, p-value = 0.43).}

@@

\revised{
\toolname only tests pipeline reproducibility in the scope of a particular dataset. However, it is very plausible 
for pipeline processes to exhibit different reproducibility behaviors when executed on different datasets. 
Therefore, only the lack of reproducibility of a pipeline process could be guaranteed from an analysis with \toolname, 
since proving reproducibility would require testing the pipeline on all possible datasets, in all possible environments, 
which is not feasible. Two elements could be considered in future work to address this issue. 
First, similar to conventional software testing, a code coverage metric could be developed to assess 
the fraction of the pipeline code involved in the tested dataset and parameters. This would quantify 
the representativity of the dataset and pipeline parameters used in the evaluation. Second, statistical 
risk models could be used to estimate the probability for a process to be reproducible, given a set of 
observations with no numerical differences. For instance, models similar to the work in~\cite{beauzamy2004methodes} 
could be leveraged for this purpose.}

@@

\revised{
\section{Availability of Source Code and Requirements}
\begin{itemize}
  \item Project name: \toolname
  \item Project home page:\url{https://github.com/big-data-lab-team/spot}
  \item Operating system: Linux
  \item Programming language: Python
  \item Other requirements: Python 3.6 or higher, \reprozip, Docker, and Boutiques
  \item License: MIT License
  \item Biotools identifier: \href{https://bio.tools/spottool}{spottool}
  \item SciCrunch ID: \href{https://scicrunch.org/scicrunch/Resources/record/nlx_144509-1/SCR_018915/resolver}{RRID:SCR\_018915}
  \item DOI: \href{https://doi.org/10.5281/zenodo.3873219}{10.5281/zenodo.3873219}

@@

\includepdf[pages=-]{supplementary-files/subject-details.pdf}
